---
title: "MATH2319_Assignment_Part1"
author: "Margaret Cuddihy (s3608125)"
date: "8 April 2019"
output: 
   html_document:
      toc: true
      toc_depth: 3
  
---
# Introduction

#### Put a description of data, goal for this report and what we will be inspecting and why.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r echo=TRUE}
 library(dplyr)
# library(tidyr)
 library(knitr)
# library(mlr)
# library(outliers)

```
# Reading and Inspecting the Data
```{r echo=TRUE}

advertising_train <- read.csv("~/Uni/Machine Learning/advertising_train.csv", header = TRUE, stringsAsFactors = FALSE)

head(advertising_train)
tail(advertising_train)
dim(advertising_train)
str(advertising_train)
names(advertising_train)
class(advertising_train)
```

## Target Feature
y: numeric revenue-related metric

## Unique Identifier
Case ID: categorical

## Descriptive Features
1. Company ID: categorical 
2. Country ID: categorical
3. Device Type: categorical (Corresponding to Desktop, Mobile, Table and unknown? Unique values = 1,2,3 and 5 )
4. Day: integer (Day 1 to Day 30 in training data)
5. DOW: categorical (day of the week)
6. Price 1 (numeric)
7. Price 2 (numeric)
8. Price 3 (numeric)
9. Ad Area (normalised between 1 and 0)
10. Ad Ratio (normalised between 1 and 0)
11. Requests (numeric)
12. Impression (numeric)
13. CPC (numeric)
14. CTR (numeric)
15. Viewability (numeric)
16. Ratio 1 (normalised between 1 and 0)
17. Ratio 2 (normalised between 1 and 0)
18. Ratio 3 (normalised between 1 and 0)
19. Ratio 4 (normalised between 1 and 0)
20. Ratio 5 (normalised between 1 and 0)

214,218 Instances

# Cleaning and Preparing Data

```{r echo=TRUE}
kable(summary(advertising_train))

advertising_train %>% group_by(ad_ratio) %>% summarize(count=n())
```

## Special Values

### Missing
```{r}
which(is.na(advertising_train)) 

# No missing values detected
```

### Special
```{r}
# Check for special values
is.special <- function(x){
  if (is.numeric(x)) !is.finite(x)
}

sapply(advertising_train, is.special)

# No special values detected
```

## Outliers

### Univariate
```{r}

```
### Multivariate
```{r}

```

# Visualisation

## Bivariate Visuals with Respect to Target Feature
```{r}

```


## Multivariate Visuals with Respect to Target Feature
```{r}

```

# Statistical Investigation

## Linear Correlations
```{r}

```

